---
title: "`r params$run_name`"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document:
    self_contained: true
    theme: cosmo
    highlight: null
    mathjax: null
    df_print: paged
params:
  bundle: ""
  run_name: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = TRUE)
suppressWarnings(suppressMessages({
  library(plotly); library(DT); library(crosstalk); library(ggplot2); library(htmltools)
}))
B <- readRDS(params$bundle)
cells <- B$cells
PAL <- {
  cts <- sort(unique(cells$cell_type))
  base <- c("#4e79a7","#f28e2b","#e15759","#76b7b2","#59a14f","#edc948","#b07aa1",
            "#ff9da7","#9c755f","#bab0ac","#86bcb6","#d37295","#fabfd2","#8cd17d","#499894")
  setNames(rep(base, length.out = length(cts)), cts)
}
n_samp <- length(B$samples)
fig_h  <- max(360, 300 * ceiling(n_samp / 3))
```

<style>
.cardrow{display:flex;flex-wrap:wrap;gap:12px;margin:14px 0}
.card{border:1px solid #e3e3e3;border-radius:10px;padding:12px 16px;min-width:150px;
      box-shadow:0 1px 2px rgba(0,0,0,.05)}
.card .v{font-size:1.6em;font-weight:600;line-height:1.1}
.card .l{color:#777;font-size:.8em;text-transform:uppercase;letter-spacing:.04em}
.card .s{color:#999;font-size:.8em}
.flag{color:#e15759;font-weight:600}
h1,h2{border-bottom:1px solid #eee;padding-bottom:4px}
body{max-width:1180px}
</style>

Interactive run report — `r B$n_cells_total` cells, `r n_samp` samples. Generated `r B$generated`.
Plots are interactive: hover for values, drag to zoom, click a legend entry to toggle that cell type across every panel. Source: `integrated_annotated.rds` (reduction `r B$umap_name`) + QC tables. No data recomputed.

# Overview

```{r cards, results='asis'}
S <- B$summary
fmt <- function(x, d=0) ifelse(is.na(x), "—", formatC(round(as.numeric(x), d), big.mark=",", format="f", digits=d))
cards <- lapply(seq_len(nrow(S)), function(i) {
  r <- S[i, ]
  low <- (!is.na(r$Pct_retained) && r$Pct_retained < 70) ||
         (!is.na(r$Median_genes) && r$Median_genes < 500)
  sprintf('<div class="card"><div class="l">%s%s</div><div class="v">%s</div>
           <div class="s">%s%% retained · %s med genes · %s types</div></div>',
          r$Sample, ifelse(low, ' <span class="flag">⚠ low QC</span>', ''),
          fmt(r$Cells), fmt(r$Pct_retained,1), fmt(r$Median_genes), fmt(r$N_cell_types))
})
cat('<div class="cardrow">', paste(unlist(cards), collapse=""), '</div>')
```

```{r summary_table}
S2 <- B$summary
names(S2) <- c("Sample","Cells","Median genes","Median UMI","Median %MT","% retained","Doublet %","Cell types","Top type")
datatable(S2, rownames = FALSE, class = "compact stripe hover",
          options = list(dom = "t", paging = FALSE, ordering = TRUE)) |>
  formatRound(c("Median %MT","% retained","Doublet %"), 1)
```

# Sample comparison {.tabset .tabset-pills}

One metric, every sample, side by side — in the integrated coordinate space, so a depleted
population appears as the same hole across panels. For runs with many samples, use **Pick samples**.

## Cell type (UMAP)

```{r umap_facet}
p <- ggplot(cells, aes(UMAP_1, UMAP_2, color = cell_type,
                       text = paste0(cell_type, "<br>", sample))) +
  geom_point(size = 0.35, alpha = 0.7) +
  scale_color_manual(values = PAL) +
  facet_wrap(~ sample, ncol = min(3, n_samp)) +
  theme_minimal(base_size = 11) +
  theme(axis.text = element_blank(), panel.grid.minor = element_blank()) +
  guides(color = guide_legend(override.aes = list(size = 3)))
ggplotly(p, tooltip = "text", height = fig_h) |>
  toWebGL() |> layout(legend = list(title = list(text = "Cell type")))
```

## Pick samples (interactive)

Filter which samples are shown. Useful when a run has more than a handful — choose the
2–4 you want to compare, then read the same coordinate space.

```{r umap_crosstalk}
sd <- SharedData$new(cells, key = ~sample, group = "samp")
sel <- filter_select("xf_sample", "Samples", sd, ~sample)
pp <- plot_ly(sd, x = ~UMAP_1, y = ~UMAP_2, color = ~cell_type, colors = PAL,
              type = "scattergl", mode = "markers",
              marker = list(size = 4, opacity = 0.6),
              text = ~paste0(cell_type, "<br>", sample), hoverinfo = "text",
              height = 560) |>
  layout(legend = list(title = list(text = "Cell type")),
         xaxis = list(title = "UMAP 1"), yaxis = list(title = "UMAP 2"))
bscols(widths = c(3, 9), sel, pp)
```

## QC

```{r qc_violin}
long <- do.call(rbind, lapply(c("nFeature_RNA","nCount_RNA","percent.mt"), function(m)
  data.frame(sample = cells$sample, metric = m, value = cells[[m]])))
long$metric <- factor(long$metric, levels = c("nFeature_RNA","nCount_RNA","percent.mt"),
                      labels = c("Genes / cell","UMIs / cell","% mito"))
gv <- ggplot(long, aes(sample, value, fill = sample)) +
  geom_violin(scale = "width", linewidth = 0.2) +
  facet_wrap(~ metric, scales = "free_y", ncol = 3) +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "none",
        axis.title = element_blank())
ggplotly(gv, height = 420)
```

## Doublets

```{r doublets}
if (all(is.na(cells$doublet_score))) {
  cat("Doublet scores not stored in this object; see the per-sample doublet PDFs.")
} else {
  gd <- ggplot(cells, aes(sample, doublet_score, fill = sample)) +
    geom_violin(scale = "width", linewidth = 0.2) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "none", axis.title.x = element_blank()) +
    labs(y = "scDblFinder score")
  ggplotly(gd, height = 380)
}
```

# Cell proportions {.tabset .tabset-pills}

## Composition

```{r prop_bar}
pl <- B$prop_long
pb <- plot_ly(pl, x = ~sample, y = ~pct, color = ~cell_type, colors = PAL,
              type = "bar", text = ~paste0(cell_type, ": ", round(pct,1), "%"),
              hoverinfo = "text", height = 460) |>
  layout(barmode = "stack", yaxis = list(title = "% of sample"),
         xaxis = list(title = ""), legend = list(title = list(text = "Cell type")))
pb
```

## Change (delta)

```{r prop_delta}
if (is.null(B$delta)) {
  cat("The pre→post delta table is shown for two-sample runs. This run has",
      length(B$samples), "samples — use the composition bars above to compare.")
} else {
  d <- B$delta; ss <- attr(d, "samples")
  dd <- datatable(d, rownames = FALSE,
                  colnames = c("Cell type", paste(ss[1], "%"), paste(ss[2], "%"), "Δ pts"),
                  class = "compact stripe hover",
                  options = list(dom = "t", paging = FALSE, order = list(list(3, "asc")))) |>
    formatStyle("delta_pts",
                color = styleInterval(c(-5, 5), c("#c0392b", "#555", "#1e8449")),
                fontWeight = "bold")
  big <- d[which.max(abs(d$delta_pts)), ]
  htmltools::tagList(
    htmltools::p(htmltools::strong("Read: "),
      sprintf("%s moves from %.1f%% (%s) to %.1f%% (%s), a change of %+.1f points — the largest shift.",
              big$cell_type, big[[2]], ss[1], big[[3]], ss[2], big$delta_pts)),
    dd)
}
```

---
<small>Generated by `pipeline/08b_html_report.R` (R Markdown + plotly). Re-render any run with
`bash pipeline/build_report.sh <run_dir>`.</small>
