---
title: "`r params$run_name`"
output:
  html_document:
    self_contained: true
    theme: cosmo
    highlight: null
    mathjax: null
    df_print: paged
params:
  bundle: ""
  run_name: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = TRUE,
                      dev = "png", dpi = 120)
suppressWarnings(suppressMessages({
  library(plotly); library(DT); library(crosstalk); library(ggplot2); library(htmltools)
}))
B <- readRDS(params$bundle)
cells <- B$cells
# violins only need enough points for the shape; downsample so ggplotly does not
# serialize ~80k cells per widget (that ballooned the file on many-sample runs)
vcells <- {
  set.seed(1)
  idx <- unlist(lapply(split(seq_len(nrow(cells)), cells$sample),
                       function(ix) if (length(ix) > 600) sample(ix, 600) else ix), use.names = FALSE)
  cells[sort(idx), , drop = FALSE]
}
PAL <- {
  cts  <- sort(unique(cells$cell_type))
  base <- c("#4e79a7","#f28e2b","#e15759","#76b7b2","#59a14f","#edc948","#b07aa1",
            "#ff9da7","#9c755f","#bab0ac","#86bcb6","#d37295","#fabfd2","#8cd17d","#499894")
  if (!is.null(B$cell_colors)) {            # canonical CELLTYPE_COLORS from the driver
    p <- unname(B$cell_colors[cts]); miss <- is.na(p)
    if (any(miss)) p[miss] <- rep(base, length.out = sum(miss))
    setNames(p, cts)
  } else setNames(rep(base, length.out = length(cts)), cts)
}
n_samp   <- length(B$samples)
# On many-sample runs the live ggplotly violins serialize ~1000-point density
# polygons per (sample x metric) trace (42 traces => ~5 MB), which is what blocked
# the file from opening. For those runs render the violins as static PNGs instead.
STATIC <- isTRUE(B$umap_static) && !is.null(B$umap_panels)
to_uri <- function(g, w, h, dpi = 72) {
  tf <- tempfile(fileext = ".png")
  ggsave(tf, plot = g, width = w, height = h, dpi = dpi, bg = "white", limitsize = FALSE)
  knitr::image_uri(tf)
}
fac_ncol <- min(3, n_samp)
fac_nrow <- ceiling(n_samp / fac_ncol)
fig_h    <- max(420, 380 * fac_nrow)
fig_w    <- max(7, 4.2 * fac_ncol)
```

```{css, echo=FALSE}
:root{
  --bg:#f4f6f9; --panel:#fff; --ink:#1f2733; --muted:#667085;
  --line:#e4e8ee; --rail:#1b2230; --rail-ink:#cdd5e0; --accent:#3b82f6;
  --good:#12a150; --warn:#e08a00; --bad:#d23f3f;
  --shadow:0 1px 2px rgba(16,24,40,.06),0 1px 3px rgba(16,24,40,.10);
}
*{box-sizing:border-box}
/* clicked nav anchors land below the fixed 52px top bar instead of behind it */
html{scroll-padding-top:68px}
body{font-family:-apple-system,Segoe UI,Roboto,Helvetica,Arial,sans-serif;
     color:var(--ink);background:var(--bg);font-size:14px;line-height:1.5}
/* hide R Markdown's default title/date header — the title lives in the top bar */
#header,h1.title,.title{display:none}

/* full-width fixed top bar */
.report-top{position:fixed;top:0;left:0;right:0;height:52px;z-index:100;display:flex;
  align-items:center;gap:14px;background:var(--rail);color:#fff;padding:0 18px;box-shadow:var(--shadow)}
.report-top .title{font-weight:600;font-size:15px;letter-spacing:.01em}
.report-top .chip{background:#2b3548;color:var(--rail-ink);border-radius:999px;padding:3px 11px;font-size:12px}
.report-top .meta{margin-left:auto;color:#9aa6b5;font-size:12px;display:flex;gap:16px;align-items:center}
.report-top .meta b{color:#fff;font-weight:600}
/* nav toggle lives at the top-right of the nav rail; when collapsed it parks at the page edge */
.navtoggle{position:fixed;top:59px;left:170px;z-index:96;width:30px;height:28px;
  background:#fff;border:1px solid var(--line);color:var(--muted);border-radius:7px;
  font-size:14px;line-height:1;cursor:pointer;box-shadow:var(--shadow);transition:left .15s}
.navtoggle:hover{color:var(--accent);border-color:var(--accent)}
body.nav-collapsed .navtoggle{left:10px}

/* fixed left navigation rail (a panel with a right border, not a card) */
.report-nav{position:fixed;top:52px;left:0;bottom:0;width:208px;background:#fff;
  border-right:1px solid var(--line);padding:10px 0;overflow:auto;z-index:90}
.report-nav a{display:block;padding:9px 18px;color:var(--muted);text-decoration:none;font-size:13px;
  border-left:3px solid transparent;cursor:pointer}
.report-nav a:hover{background:#f7f9fc;color:var(--ink)}
.report-nav a.active{color:var(--ink);font-weight:600;border-left-color:var(--accent);background:#f2f6ff}
.report-nav .grouplabel{font-size:11px;text-transform:uppercase;letter-spacing:.04em;color:#9aa6b5;
  padding:14px 18px 4px}

/* main column clears the fixed top bar + nav; collapsing the nav frees the whole page */
.main-container{margin:52px 0 0 208px !important;max-width:1080px;padding:10px 30px 70px;transition:margin .15s}
body.nav-collapsed .report-nav{display:none}
body.nav-collapsed .main-container{margin-left:auto !important;margin-right:auto}

h1{font-size:20px;border-bottom:1px solid var(--line);padding-bottom:6px;margin-top:30px}
h2{font-size:15px;border:0;margin-top:16px}
h1:empty{display:none}  /* the hidden tab-closing header after the galleries */

/* sample cards: compact, label/value stat rows (mockup style) */
.cardgrid{display:grid;grid-template-columns:repeat(auto-fill,minmax(196px,1fr));gap:12px;
  margin:10px 0 22px;max-width:880px}
.scard{background:#fff;border:1px solid var(--line);border-radius:10px;padding:12px 14px;box-shadow:var(--shadow)}
.scard h3{margin:0 0 9px;font-size:13.5px;display:flex;align-items:center;gap:7px;font-weight:600}
.scard .tag{font-size:10.5px;color:#fff;background:var(--bad);border-radius:999px;padding:1px 7px;font-weight:600}
.scard .stat{display:flex;justify-content:space-between;padding:2.5px 0;font-size:12.5px;color:var(--muted)}
.scard .stat b{color:var(--ink);font-variant-numeric:tabular-nums;font-weight:600}

/* accent tab pills + tables + read line */
.nav-pills>li.active>a,.nav-pills>li.active>a:hover,.nav-pills>li.active>a:focus{background:var(--accent)}
.nav-pills>li>a{color:var(--muted)}
table.dataTable thead th{background:#fafbfd;color:var(--muted);font-weight:600;border-bottom:1px solid var(--line)}
/* overview summary rows deep-link to the per-sample plots */
table.clickrows tbody tr{cursor:pointer}
table.clickrows tbody tr:hover td{background:#f2f6ff}
table.clickrows caption{caption-side:top;text-align:left;color:var(--muted);font-size:12.5px;padding:0 0 7px}
.read-line{background:#f2f6ff;border-left:3px solid var(--accent);border-radius:6px;
  padding:9px 14px;margin:4px 0 14px;font-size:13.5px}
hr{border-top:1px solid var(--line)}
small{color:var(--muted)}

/* sample selector: rounded toggle chips (restyled crosstalk checkbox group), no ticks, no search */
.crosstalk-input{margin:0 0 14px}
.crosstalk-input .control-label{font-size:11px;text-transform:uppercase;letter-spacing:.05em;
  color:#9aa6b5;font-weight:600;margin-bottom:8px;display:block}
.crosstalk-input-checkboxgroup .crosstalk-options-group{display:flex;flex-wrap:wrap;gap:8px}
.crosstalk-input-checkboxgroup .checkbox{margin:0}
.crosstalk-input-checkboxgroup .checkbox input{position:absolute;opacity:0;width:0;height:0;pointer-events:none}
.crosstalk-input-checkboxgroup .checkbox label{display:inline-flex;align-items:center;border:1px solid var(--line);
  border-radius:999px;padding:6px 15px;font-size:13px;background:#fff;color:var(--muted);cursor:pointer;
  font-weight:500;user-select:none}
.crosstalk-input-checkboxgroup .checkbox label:hover{border-color:#c4ccd8;color:var(--ink)}
.crosstalk-input-checkboxgroup .checkbox label:has(input:checked){border-color:var(--accent);
  background:#f2f6ff;color:var(--ink);box-shadow:inset 0 0 0 1px var(--accent)}

/* sample-comparison: chip selector row + UMAP grid (max 3/row) + one shared legend on the side */
.selrow{display:flex;flex-wrap:wrap;gap:8px;align-items:center;margin:10px 0 16px}
.selrow .sellabel{font-size:11px;text-transform:uppercase;letter-spacing:.05em;color:#9aa6b5;
  font-weight:600;margin-right:4px}
.selchip{background:#fff;border:1px solid var(--line);border-radius:999px;padding:6px 15px;font-size:13px;
  cursor:pointer;color:var(--muted);font-weight:500}
.selchip:hover{border-color:#c4ccd8;color:var(--ink)}
.selchip.on{border-color:var(--accent);background:#f2f6ff;color:var(--ink);box-shadow:inset 0 0 0 1px var(--accent)}
.umapwrap{display:flex;gap:16px;align-items:flex-start}
.umapgrid{display:grid;grid-template-columns:repeat(2,1fr);gap:16px;flex:1;min-width:0}
/* one visible panel (single-sample run, or all but one chip deselected) spans full width */
.umapgrid.solo{grid-template-columns:1fr}
.umapcell{border:1px solid var(--line);border-radius:10px;background:#fff;padding:6px 6px 2px;box-shadow:var(--shadow);min-width:0}
.umaptitle{font-weight:600;font-size:12.5px;padding:3px 6px;color:var(--ink)}
.shlegend{width:158px;flex:none;border:1px solid var(--line);border-radius:10px;background:#fff;
  padding:10px 12px;box-shadow:var(--shadow);position:sticky;top:64px}
.shlegend .lgtitle{font-size:11px;text-transform:uppercase;letter-spacing:.04em;color:#9aa6b5;
  font-weight:600;margin-bottom:8px}
.legitem{display:flex;align-items:center;gap:8px;padding:4px 2px;font-size:12.5px;cursor:pointer;color:var(--ink)}
.legitem:hover{color:var(--accent)}
.legitem.off{opacity:.38}
.legitem.static{cursor:default}
.legitem.static:hover{color:var(--ink)}
.legsw{width:12px;height:12px;border-radius:3px;flex:none}
.muted-note{color:var(--muted);font-size:12.5px;margin:0 0 12px;max-width:760px}

/* clickable sample cards (deep-link into the per-sample gallery) */
.scardlink{text-decoration:none;color:inherit;display:block}
.scardlink:hover .scard{border-color:var(--accent);box-shadow:0 2px 7px rgba(59,130,246,.16)}

/* one- or two-line grey description printed under a section title */
.plot-desc{color:var(--muted);font-size:12.5px;line-height:1.45;margin:2px 0 14px;max-width:820px}

/* static gallery figures — ONE consistent, bounded width for every per-stage and
   per-sample plot (like the comparison panels; not full-bleed). height:auto keeps each
   plot's own proportions; sources are rendered a touch wider so this never upscales. */
.galfig{margin:10px 0 20px}
.galfig figcaption{font-size:12.5px;color:var(--muted);font-weight:600;margin-bottom:5px}
.galfig img{width:100%;max-width:680px;height:auto;border:1px solid var(--line);border-radius:8px;background:#fff}

/* per-sample plots in a 2-up grid so the section is not an endless single column */
.galgrid{display:grid;grid-template-columns:repeat(2,minmax(0,1fr));gap:4px 18px;align-items:start}
.galfig figcaption{color:var(--ink);font-size:13px}   /* darker caption vs the grey description */

/* stronger per-section headers in the galleries so samples/stages don't blur together */
#per-sample h2,#galleries h2{margin-top:28px;padding:6px 11px;background:#f2f6ff;
  border-left:3px solid var(--accent);border-radius:6px;font-size:15.5px;color:var(--ink)}

/* WCAG-AA contrast: darken the small on-white labels (the on-dark top-bar meta is left alone) */
.report-nav .grouplabel,.crosstalk-input .control-label,.selrow .sellabel,.shlegend .lgtitle{color:#6b7686}
.scard .tag{background:#c0392b}

/* responsive: below 820px the fixed 208px rail becomes an off-canvas drawer and every
   multi-column grid collapses to one column (the desktop layout otherwise squeezes the
   main content to a sliver on tablets/phones and overflows wide tables) */
@media (max-width: 820px){
  .report-nav{transform:translateX(-100%);transition:transform .2s ease;width:240px;
    box-shadow:2px 0 12px rgba(16,24,40,.18)}
  body.nav-open .report-nav{transform:none}
  .navtoggle{left:10px;top:58px}
  .main-container{margin-left:0 !important;margin-right:auto;padding:10px 16px 60px}
  .report-top .meta{display:none}
  .report-top .title{font-size:14px}
  .umapwrap{flex-direction:column}
  .umapgrid{grid-template-columns:1fr}
  .shlegend{position:static;width:auto;top:auto}
  .galgrid{grid-template-columns:1fr}
  .galfig img{max-width:100%}
}
```

```{r shell, results='asis'}
top <- sprintf('<div class="report-top"><span class="title">%s</span><span class="chip">Interactive run report</span><span class="meta"><span><b>%s</b> cells</span><span><b>%s</b> samples</span><span>%s</span></span></div><button class="navtoggle" title="Show / hide the navigation rail" onclick="document.body.classList.toggle(matchMedia(\'(max-width:820px)\').matches?\'nav-open\':\'nav-collapsed\')">&#9776;</button>',
  htmltools::htmlEscape(B$run_name), format(B$n_cells_total, big.mark=","), n_samp,
  paste(format(Sys.Date(), "%d"), month.abb[as.integer(format(Sys.Date(), "%m"))], format(Sys.Date(), "%Y")))
slug <- function(x) gsub("(^-|-$)", "", gsub("[^a-z0-9]+", "-", tolower(x)))
gal  <- !isTRUE(B$lite) && !is.null(B$galleries)
stage_links <- if (gal && length(B$galleries$stage)) paste0('<div class="grouplabel">Per stage</div>',
  paste(sprintf('<a href="#stage-%s">%s</a>', slug(names(B$galleries$stage)),
                htmltools::htmlEscape(names(B$galleries$stage))), collapse="")) else ''
sample_links <- if (gal && length(B$galleries$by_sample)) paste0('<div class="grouplabel">Per sample</div>',
  paste(sprintf('<a href="#sample-%s">%s</a>', slug(names(B$galleries$by_sample)),
                htmltools::htmlEscape(names(B$galleries$by_sample))), collapse="")) else ''
nav <- paste0('<nav class="report-nav">',
  '<div class="grouplabel">Run</div><a href="#overview">Overview</a>',
  '<div class="grouplabel">Compare</div>',
  '<a href="#sample-comparison">Sample comparison</a>',
  '<a href="#cell-proportions">Cell proportions</a>',
  '<a href="#cell-type-markers">Cell-type markers</a>',
  stage_links, sample_links,
  '<div class="grouplabel">About</div><a href="#versions">Versions</a>',
  '</nav>')
cat(top, nav)
```

Hover any plot for values, drag to zoom, and click a cell type in a legend to toggle it across every
panel — that toggle *is* the depletion view. Source: `integrated_annotated.rds` (reduction
`r B$umap_name`) + the QC tables. No data is recomputed.

# Overview

```{r cards, results='asis'}
S <- B$summary
fmt <- function(x, d=0) ifelse(is.na(x), "—", formatC(round(as.numeric(x), d), big.mark=",", format="f", digits=d))
row <- function(l, v) sprintf('<div class="stat"><span>%s</span><b>%s</b></div>', l, v)
cards <- vapply(seq_len(nrow(S)), function(i) {
  r <- S[i, ]
  low <- (!is.na(r$Pct_retained) && r$Pct_retained < 70) ||
         (!is.na(r$Median_genes) && r$Median_genes < 500)
  tag <- if (low) ' <span class="tag">low QC</span>' else ''
  stats <- paste0(
    row("Cells", fmt(r$Cells)),
    row("% retained", paste0(fmt(r$Pct_retained,1), "%")),
    row("Median genes", fmt(r$Median_genes)),
    row("Median UMI", fmt(r$Median_UMI)),
    row("% MT", paste0(fmt(r$Median_pct_MT,1), "%")),
    if (!is.null(r$Doublet_pct) && !is.na(r$Doublet_pct)) row("Doublets", paste0(fmt(r$Doublet_pct,1), "%")) else "",
    if (!is.null(r$HVG) && !is.na(r$HVG)) row("HVG", fmt(r$HVG)) else "",
    row("Cell types", fmt(r$N_cell_types)))
  card <- sprintf('<div class="scard"><h3>%s%s</h3>%s</div>', htmltools::htmlEscape(r$Sample), tag, stats)
  slug <- function(x) gsub("(^-|-$)", "", gsub("[^a-z0-9]+", "-", tolower(x)))
  has_gal <- !isTRUE(B$lite) && !is.null(B$galleries) && r$Sample %in% names(B$galleries$by_sample)
  if (has_gal) sprintf('<a class="scardlink" href="#sample-%s">%s</a>', slug(r$Sample), card) else card
}, character(1))
cat('<div class="cardgrid">', paste(cards, collapse=""), '</div>')
```

```{r summary_table}
S2 <- B$summary
nm  <- c(Sample="Sample", Cells="Cells", Median_genes="Median genes", Median_UMI="Median UMI",
         Median_pct_MT="Median %MT", Median_pct_ribo="Median %ribo", Pct_retained="% retained",
         Doublet_pct="Doublet %", HVG="HVG", N_cell_types="Cell types", Top_type="Top type")
names(S2) <- ifelse(names(S2) %in% names(nm), nm[names(S2)], names(S2))
roundable <- intersect(c("Median %MT","Median %ribo","% retained","Doublet %"), names(S2))
# clickable rows: jump to that sample's per-sample plots (same slug as the section ids)
gal_ok <- !isTRUE(B$lite) && !is.null(B$galleries) && length(B$galleries$by_sample)
cb <- DT::JS(
  "var slugify=function(s){return String(s).toLowerCase().replace(/[^a-z0-9]+/g,'-').replace(/^-|-$/g,'');};",
  "table.on('click','tbody tr',function(){",
  "  var d=table.row(this).data(); if(!d) return;",
  "  var id='sample-'+slugify(d[0]);",
  "  if(document.getElementById(id)){location.hash=id;}",
  "});")
datatable(S2, rownames = FALSE,
          class = if (gal_ok) "compact stripe hover clickrows" else "compact stripe hover",
          caption = if (gal_ok) "Click a row to open that sample's detailed plots below." else NULL,
          selection = "single",
          callback = if (gal_ok) cb else NULL,
          options = list(dom = "t", paging = FALSE, ordering = TRUE, scrollX = TRUE)) |>
  formatRound(roundable, 1)
```

# Sample comparison {.tabset .tabset-pills}

Pick a metric below, then choose which samples to compare with the chips. Every panel shares the
integrated coordinate space, so a depleted population shows up as the same hole across samples.

## UMAP

```{r umap_grid}
slugv  <- function(x) gsub("(^-|-$)", "", gsub("[^a-z0-9]+", "-", tolower(x)))
cts    <- sort(unique(cells$cell_type))
static <- isTRUE(B$umap_static) && !is.null(B$umap_panels)

if (static) {
  # many-sample runs: light static PNG panels (no 14x WebGL contexts / 16 MB of JSON)
  panels <- lapply(B$samples, function(s)
    htmltools::div(class = "umapcell", id = paste0("umapwrap-", slugv(s)),
      htmltools::div(class = "umaptitle", s),
      htmltools::tags$img(src = B$umap_panels[[s]], style = "width:100%;display:block;border-radius:6px")))
} else {
  mk_umap <- function(s) {
    dd <- cells[cells$sample == s, , drop = FALSE]
    nc <- B$summary$Cells[match(s, B$summary$Sample)]; if (is.na(nc)) nc <- nrow(dd)
    plot_ly(dd, x = ~UMAP_1, y = ~UMAP_2, color = ~cell_type, colors = PAL,
            type = "scattergl", mode = "markers",
            marker = list(size = 4, opacity = 0.8),
            text = ~cell_type, hoverinfo = "text", showlegend = FALSE, height = 360) |>
      layout(title = list(text = sprintf("%s  (%s cells)", s, format(nc, big.mark = ",")),
                          x = 0.03, y = 0.97, font = list(size = 13)),
             xaxis = list(title = "UMAP 1", zeroline = FALSE, showticklabels = FALSE),
             yaxis = list(title = "UMAP 2", zeroline = FALSE, showticklabels = FALSE),
             margin = list(l = 42, r = 8, t = 30, b = 36)) |>
      toWebGL()
  }
  panels <- lapply(B$samples, function(s)
    htmltools::div(class = "umapcell", id = paste0("umapwrap-", slugv(s)),
                   htmltools::div(class = "umaptitle", s), mk_umap(s)))
}
chips <- lapply(B$samples, function(s)
  htmltools::tags$button(class = "selchip on", `data-sample` = slugv(s), s))
legitems <- lapply(cts, function(ct)
  htmltools::tags$div(class = if (static) "legitem static" else "legitem", `data-ct` = ct,
    htmltools::tags$span(class = "legsw", style = sprintf("background:%s", PAL[[ct]])),
    htmltools::tags$span(ct)))
js_solo <- "var _wasSolo=null;function umResize(c,solo){var w=c.querySelector('.html-widget'),p=c.querySelector('.js-plotly-plot');if(w)w.style.height=(solo?700:360)+'px';if(p&&window.Plotly){try{Plotly.Plots.resize(p);}catch(e){}}}function umSolo(){var g=document.querySelector('.umapgrid');if(!g)return;var n=0;Array.prototype.forEach.call(g.children,function(c){if(c.style.display!=='none')n++;});var solo=n===1;g.classList.toggle('solo',solo);if(solo===_wasSolo)return;_wasSolo=solo;requestAnimationFrame(function(){Array.prototype.forEach.call(g.children,function(c){if(c.style.display!=='none')umResize(c,solo);});});}"
js_chip <- "document.querySelectorAll('.selrow .selchip').forEach(function(b){b.addEventListener('click',function(){var on=b.classList.toggle('on');var w=document.getElementById('umapwrap-'+b.getAttribute('data-sample'));if(w){w.style.display=on?'':'none';if(on)requestAnimationFrame(function(){var p=w.querySelector('.js-plotly-plot');if(p&&window.Plotly){try{Plotly.Plots.resize(p);}catch(e){}}});}umSolo();});});"
js_leg  <- if (static) "" else "document.querySelectorAll('.shlegend .legitem').forEach(function(it){it.addEventListener('click',function(){var off=it.classList.toggle('off'),name=it.getAttribute('data-ct');document.querySelectorAll('.umapgrid .js-plotly-plot').forEach(function(gd){if(!gd.data) return;var idx=[];gd.data.forEach(function(t,i){if(t.name===name) idx.push(i);});if(idx.length&&window.Plotly) Plotly.restyle(gd,{visible:off?'legendonly':true},idx);});});});"
note <- if (static) htmltools::tags$p(class = "muted-note",
  sprintf("%d samples: panels are high-resolution static snapshots (live zoom + legend toggle kick in automatically for runs of 8 samples or fewer). Use the chips to show or hide samples.", length(B$samples))) else NULL
htmltools::browsable(htmltools::tagList(
  htmltools::div(class = "selrow", htmltools::tags$span(class = "sellabel", "Samples"), chips),
  note,
  htmltools::div(class = "umapwrap",
    htmltools::div(class = "umapgrid", panels),
    htmltools::div(class = "shlegend",
      htmltools::div(class = "lgtitle", "Cell type"), legitems)),
  htmltools::tags$script(htmltools::HTML(sprintf("(function(){%s%s%s umSolo();})();",
    js_solo, js_chip, js_leg)))
))
```

Click a sample chip to add or drop its panel. Click a cell type in the shared legend to toggle it
across every UMAP at once (e.g. isolate CD14+ Mono pre vs post to read the sort).

## QC

<p class="plot-desc">Distribution of genes detected, UMIs and mitochondrial % per cell, one violin per sample
(PDF style — no scatter dots). Hover for the median and quartiles; compare spreads to spot
low-complexity or high-MT samples. Drag to zoom.</p>

```{r qc_violin}
# Use ggplot geom_violin(scale="width") — the SAME engine as the pipeline's Seurat
# VlnPlot (01_load_qc.R) — so the violin shapes match Overall_Report.pdf exactly.
# trim=TRUE clips the density to the data range; no jitter points (PDF style).
long <- do.call(rbind, lapply(c("nFeature_RNA","nCount_RNA","percent.mt"), function(m)
  data.frame(sample = vcells$sample, metric = m, value = vcells[[m]])))
long$metric <- factor(long$metric, levels = c("nFeature_RNA","nCount_RNA","percent.mt"),
                      labels = c("Genes / cell","UMIs / cell","% mito"))
gv <- ggplot(long, aes(sample, value, fill = sample)) +
  geom_violin(scale = "width", trim = TRUE, linewidth = 0.2) +
  facet_wrap(~ metric, scales = "free_y", ncol = 3) +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position = "none",
        axis.title = element_blank(), strip.text = element_text(face = "bold"))
ggplotly(gv, height = 430, tooltip = "y")
```

## Doublets

<p class="plot-desc">Residual scDblFinder scores of the <i>retained</i> cells per sample (interactive violin, no dots) —
called doublets were already removed before integration, so this shows how cleanly each sample
separated. For the per-sample doublet calls and counts, see the Doublets panels in Per-sample detail.</p>

```{r doublets}
if (all(is.na(cells$doublet_score))) {
  cat("Doublet scores are not stored in this object; see the per-sample doublet plots below.")
} else {
  gd <- ggplot(vcells, aes(sample, doublet_score, fill = sample)) +
    geom_violin(scale = "width", trim = TRUE, linewidth = 0.2) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "none", axis.title.x = element_blank(),
          axis.text.x = element_text(angle = 30, hjust = 1)) +
    labs(y = "scDblFinder score")
  ggplotly(gd, height = 390, tooltip = "y")
}
```

# Cell proportions {.tabset .tabset-pills}

## Composition

```{r prop_bar}
pl <- B$prop_long
pb <- plot_ly(pl, x = ~sample, y = ~pct, color = ~cell_type, colors = PAL,
              type = "bar", text = ~paste0(cell_type, ": ", round(pct,1), "%"),
              hoverinfo = "text", height = 470) |>
  layout(barmode = "stack", yaxis = list(title = "% of sample"),
         xaxis = list(title = ""), legend = list(title = list(text = "Cell type")),
         updatemenus = list(list(type = "buttons", direction = "right",
           x = 0, y = 1.08, xanchor = "left", showactive = FALSE, pad = list(t = 2),
           buttons = list(
             list(label = "Show all", method = "restyle", args = list("visible", TRUE)),
             list(label = "Hide all", method = "restyle", args = list("visible", "legendonly"))))))
pb
```

Click a cell type in the legend to drop it from the bars (or **Hide all**, then click one back on).
The myeloid fraction here doubles as a contamination read: a high CD14+ Mono / granulocyte share in a
lymphocyte-sorted sample flags carry-over.

## Cell numbers

```{r count_bar}
cl <- B$counts_long
if (is.null(cl) || !nrow(cl)) {
  cat("Absolute cell counts are unavailable for this run.")
} else {
  cb <- plot_ly(cl, x = ~sample, y = ~n, color = ~cell_type, colors = PAL,
                type = "bar", text = ~paste0(cell_type, ": ", format(n, big.mark = ","), " cells"),
                hoverinfo = "text", height = 470) |>
    layout(barmode = "stack", yaxis = list(title = "cells"),
           xaxis = list(title = ""), legend = list(title = list(text = "Cell type")),
           updatemenus = list(list(type = "buttons", direction = "right",
             x = 0, y = 1.08, xanchor = "left", showactive = FALSE, pad = list(t = 2),
             buttons = list(
               list(label = "Show all", method = "restyle", args = list("visible", TRUE)),
               list(label = "Hide all", method = "restyle", args = list("visible", "legendonly"))))))
  cb
}
```

Same stacked bars as Composition, but **absolute cell counts** per sample (one column each). Tall bars are
deeply sequenced samples; compare against the proportions tab to separate "more cells" from "different mix".

## Change (delta)

```{r prop_delta}
if (is.null(B$delta)) {
  cat("The pre→post delta table is shown for two-sample runs. This run has",
      length(B$samples), "samples — use the composition bars above to compare.")
} else {
  d <- B$delta; ss <- attr(d, "samples")
  big <- d[which.max(abs(d$delta_pts)), ]
  read <- sprintf('<div class="read-line"><b>Read:</b> %s moves from %.1f%% (%s) to %.1f%% (%s), a change of %+.1f points — the largest shift.</div>',
                  htmltools::htmlEscape(big$cell_type), big[[2]], ss[1], big[[3]], ss[2], big$delta_pts)
  dd <- datatable(d, rownames = FALSE,
                  colnames = c("Cell type", paste(ss[1], "%"), paste(ss[2], "%"), "Δ pts"),
                  class = "compact stripe hover",
                  options = list(dom = "t", paging = FALSE, scrollX = TRUE, order = list(list(3, "asc")))) |>
    formatStyle("delta_pts",
                color = styleInterval(c(-5, 5), c("#c0392b", "#555", "#1e8449")),
                fontWeight = "bold")
  htmltools::tagList(htmltools::HTML(read), dd)
}
```

# Cell-type markers

Canonical marker panel across the annotated types — dot **size** is the % of cells in that type
expressing the gene, **colour** is the scaled average expression. This is the evidence behind the
labels (e.g. CD14/LYZ/S100A8 lighting up CD14+ Mono, MS4A1/CD79A on B cells).

```{r marker_dot}
md_ <- B$markers_dot
if (is.null(md_) || !nrow(md_)) {
  cat("Marker dot plot is unavailable for this run.")
} else {
  go <- attr(md_, "gene_order"); if (is.null(go)) go <- unique(md_$gene)
  md_$gene <- factor(md_$gene, levels = go)
  md_$cell_type <- factor(md_$cell_type, levels = rev(sort(unique(md_$cell_type))))
  gp <- ggplot(md_, aes(gene, cell_type, size = pct_exp, color = avg_scaled,
                 text = paste0("<b>", gene, "</b> in ", cell_type, "<br>",
                               round(pct_exp), "% expressing<br>scaled avg ", round(avg_scaled, 2)))) +
    geom_point() +
    scale_color_viridis_c(option = "plasma", name = "scaled avg") +
    scale_size(range = c(0, 6), name = "% expressing") +
    theme_minimal(base_size = 11) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 10.5),
          axis.text.y = element_text(size = 10.5),
          axis.title = element_blank(), panel.grid.major = element_line(color = "#f0f0f0"))
  ggplotly(gp, tooltip = "text", height = max(360, 26 * nlevels(md_$cell_type) + 150)) |>
    layout(xaxis = list(automargin = TRUE), yaxis = list(automargin = TRUE))
}
```

```{r galleries, results='asis'}
# Emit real markdown headers (not raw <h1>/<h2>) so pandoc CLOSES the proportions
# {.tabset} above. Raw HTML headers left the galleries trapped inside the last (hidden)
# tab-pane, so anchor links to them could not scroll. {#id} sets the nav/card targets.
if (!isTRUE(B$lite) && !is.null(B$galleries)) {
  G <- B$galleries
  slug <- function(x) gsub("(^-|-$)", "", gsub("[^a-z0-9]+", "-", tolower(x)))
  # one- or two-line plain-English description for each plot family (cf. Overall_Report.pdf)
  DESC <- c(
    "Top DE marker heatmap" =
      "Top differentially expressed genes per cell type (rows), scaled across types (columns). Bright red marks the type where each gene is most enriched — the molecular signature behind each label.",
    "Key lineage marker expression" =
      "Canonical lineage genes projected on the shared UMAP. Each panel lights up where one marker is expressed (CD3D = T, MS4A1/CD79A = B, NKG7/GNLY = NK, CD14/LYZ = monocyte), confirming the annotation spatially.",
    "SingleR reference heatmap" =
      "SingleR correlation scores of every cell (columns) against each reference label (rows). Strong diagonal blocks indicate confident, unambiguous assignments.",
    "QC: UMIs vs genes" =
      "Each dot is a cell: total UMIs vs genes detected, coloured by mitochondrial %. The dashed box is the acceptance window — cells outside it were filtered out.",
    "QC: UMIs vs % MT" =
      "UMIs vs mitochondrial fraction per cell. The dashed line is the %MT cutoff; stressed or dying cells (high MT, low UMI) sit above it and are removed.",
    "Doublets (UMAP)" =
      "Per-sample UMAP (pre-removal, from step 02) coloured by scDblFinder call. Red doublets tend to fall between lineages or in high-UMI regions and are dropped before integration. Counts reflect the full sample, not the plotting cap.",
    "Doublet score distribution" =
      "Histogram of scDblFinder doublet scores for the sample, coloured by called class. The singlet/doublet split shows where the score threshold fell.",
    "Highly variable genes" =
      sprintf("Mean expression vs standardized variance per gene; the 10 most variable are labelled. The dashed line marks standardized variance %s — the Overview HVG count is the number of genes above it. These variable genes drive PCA and clustering.",
              if (is.null(B$hvg_var_threshold)) "1.3" else B$hvg_var_threshold))
  desc_of <- function(name) if (!is.na(DESC[name])) sprintf('<p class="plot-desc">%s</p>\n', DESC[name]) else ""
  emit_fig <- function(uri, cap = NULL) {
    # caption (one line) ABOVE the image, description BELOW it — so two figures sharing a grid
    # row have their image tops flush regardless of how long each description is
    fc <- if (is.null(cap)) "" else sprintf('<figcaption>%s</figcaption>', htmltools::htmlEscape(cap))
    desc <- if (is.null(cap)) "" else desc_of(cap)
    cat(sprintf('\n<figure class="galfig">%s<img src="%s">%s</figure>\n', fc, uri, desc))
  }
  if (length(G$stage)) {
    cat('\n# Stage galleries {#galleries}\n\n')
    cat('Regenerated from the saved objects. The SingleR panels stand in for the raw score heatmap (the pipeline does not keep the score matrix in the object).\n')
    for (t in names(G$stage)) {
      cat(sprintf('\n## %s {#stage-%s}\n\n', t, slug(t)))
      cat(desc_of(t))
      emit_fig(G$stage[[t]])
    }
  }
  if (length(G$by_sample)) {
    cat('\n# Per-sample detail {#per-sample}\n\n')
    cat('Pick a sample to load its diagnostics (so you do not scroll through every sample). The Overview cards, the summary table, and the left-nav sample links jump here too. Doublet panels appear only for samples whose doublet detection ran in this analysis; cache-loaded samples (pre-removed singlets) show the other diagnostics only.\n')
    sn <- names(G$by_sample)
    chips_ps <- paste(sprintf('<button class="selchip" data-psample="%s">%s</button>',
                              slug(sn), htmltools::htmlEscape(sn)), collapse = "")
    cat(sprintf('\n<div class="selrow psel"><span class="sellabel">Sample</span>%s</div>\n', chips_ps))
    for (s in sn) {
      cat(sprintf('\n## %s {#sample-%s}\n\n', s, slug(s)))
      cat('\n<div class="galgrid">\n')           # 2-up grid so a sample is not an endless column
      imgs <- G$by_sample[[s]]; for (t in names(imgs)) emit_fig(imgs[[t]], t)
      cat('\n</div>\n')
    }
    # show one sample at a time; reveal on chip click or on a #sample-<slug> deep-link
    cat('<script>(function(){var secs=document.querySelectorAll("#per-sample .level2");var chips=document.querySelectorAll(".psel .selchip");if(!secs.length)return;var hdr=document.getElementById("per-sample");function show(sl){secs.forEach(function(s){s.style.display=(s.id==="sample-"+sl)?"":"none";});chips.forEach(function(c){c.classList.toggle("on",c.getAttribute("data-psample")===sl);});}function goPS(sl){show(sl);if(hdr)hdr.scrollIntoView();}chips.forEach(function(c){c.addEventListener("click",function(){goPS(c.getAttribute("data-psample"));});});function fromHash(){var m=(location.hash||"").match(/^#sample-(.+)$/);if(m)goPS(m[1]);}window.addEventListener("hashchange",fromHash);if(/^#sample-/.test(location.hash)){fromHash();}else{show(chips[0].getAttribute("data-psample"));}})();</script>\n')
  }
}
```

# Versions {#versions}

<p class="plot-desc">Versions of the single-cell analysis packages used to produce this run. Report-generation
tooling (R Markdown, plotly, etc.) is intentionally excluded.</p>

```{r versions}
tv <- B$tool_versions
if (is.null(tv) || !nrow(tv)) {
  cat("Tool versions are unavailable for this run.")
} else {
  if ("Source" %in% names(tv))
    tv$Source <- sprintf('<a href="%s" target="_blank" rel="noopener">%s</a>',
                         tv$Source, sub("^https?://(www\\.)?", "", tv$Source))
  datatable(tv, rownames = FALSE, class = "compact stripe hover", escape = FALSE,
            options = list(dom = "t", paging = FALSE, ordering = FALSE, scrollX = TRUE))
}
```

# {.unlisted .unnumbered}

<small>Generated by `pipeline/08b_html_report.R` (R Markdown + plotly). Re-render any run with
`bash pipeline/build_report.sh <run_dir>`.</small>

```{r scrollspy, results='asis'}
cat('<script>
(function(){
  var links=[].slice.call(document.querySelectorAll(".report-nav a"));
  var secs=links.map(function(a){return document.querySelector(a.getAttribute("href"));});
  function onScroll(){
    var y=(window.scrollY||window.pageYOffset)+90, idx=0;
    for(var i=0;i<secs.length;i++){ if(secs[i] && secs[i].offsetTop<=y) idx=i; }
    links.forEach(function(a,i){ a.classList.toggle("active", i===idx); });
  }
  window.addEventListener("scroll", onScroll, {passive:true});
  window.addEventListener("load", onScroll); onScroll();
})();
</script>')
```
